---
output:
  md_document:
    variant: gfm
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# CaribouBC

> Caribou Population Forecasting

[![Linux build status](https://travis-ci.org/psolymos/CaribouBC.svg?branch=master)](https://travis-ci.org/psolymos/CaribouBC)
[![codecov](https://codecov.io/gh/psolymos/CaribouBC/branch/master/graph/badge.svg)](https://codecov.io/gh/psolymos/CaribouBC)

Try the [Shiny app](https://psolymos.shinyapps.io/matpen/) or run locally as 
`shiny::runGitHub("bcgov/CaribouBC", subdir = "inst/shiny/matpen")`.

[](caribou.gif)

## Installation

Stable version:

```{r install,eval=FALSE}
devtools::install_github("bcgov/CaribouBC")
```

Development version is available as this fork:

```{r install_dev,eval=FALSE}
devtools::install_github("psolymos/CaribouBC")
```

## Usage

```{r example}
library(CaribouBC)

## Predefined settings
(s1 <- caribou_settings("mat.pen"))
(s2 <- caribou_settings("pred.excl"))

## Modifying predefined settings
caribou_settings("mat.pen", c.surv.capt=0.65, pen.cap=30)

## Forecast based on settings for 75% females penned
f1 <- caribou_forecast(s1, fpen.prop = 0.75)
f2 <- caribou_forecast(s2, fpen.prop = 0.75)

## Most important results summarized
summary(f1)
summary(f2)

## Plot the results
plot(f2)
lines(f1, col = 2)
legend("topleft", col = c(1,1,2), lty = c(2,1,1),
    legend = c("No pen", "Mat pen", "Pred excl"))

## Find 'breakeven' proportion of females penned where lambda=1
(b1 <- caribou_breakeven(f1, lambda = 1))
(b2 <- caribou_breakeven(f2, lambda = 1))
f3 <- caribou_forecast(s1, fpen.prop = b1)
f4 <- caribou_forecast(s2, fpen.prop = b2)
## See that lines are truly flat
op <- par(mfrow = c(1, 2))
plot(f3, main = "Mat pen")
plot(f4, main = "Pred excl")
par(op)

## Forecast using number of penned females
## - it can be an initial number
## - or a vector of individuals for subsequent years
caribou_forecast(s1, fpen.inds = 5)
caribou_forecast(s2, fpen.inds = c(5, 0, 4, 6))
```

## Getting Help or Reporting an Issue

To report bugs/issues/feature requests, please file an [issue](https://github.com/bcgov/CaribouBC/issues/).

## How to Contribute

If you would like to contribute to the package, please see our 
[CONTRIBUTING](CONTRIBUTING.md) guidelines.

Please note that this project is released with a [Contributor Code of Conduct](CODE_OF_CONDUCT.md). By participating in this project you agree to abide by its terms.

## License

```
Copyright 2018 Province of British Columbia

Licensed under the Apache License, Version 2.0 (the &quot;License&quot;);
you may not use this file except in compliance with the License.
You may obtain a copy of the License at

http://www.apache.org/licenses/LICENSE-2.0

Unless required by applicable law or agreed to in writing, software distributed under the License is distributed on an &quot;AS IS&quot; BASIS,
WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
See the License for the specific language governing permissions and limitations under the License.
```

---
*This project was created using the [bcgovr](https://github.com/bcgov/bcgovr) package.* 
